\mytitle

### 2016-06-16 pdocs/thesis/article-introduction.tex
pdocs/article changes based on Aki's comments

Physics engine is computer software that provides an approximate simulation of physical system. 
E.g. ODE - \cite{ode}, \bullet - \cite{bullet}, Box2D - \cite{box2d} are open source physics engines that are  widely used in 
areas where fast solutions are required. Typical industries are games and film productions.

Computational methods used in physics engines are divided to modules that handle collision detection and modules that handle simulation.
In this work focus is on solving new types of contraints between rigid objects.
Detailed description of various components can be found in e.g. \cite{erleben.thesis}.

Plasticity is not typically taken into account in gaming solutions. 
Breaking of various objects typically takes place based on collision or impulse.
Theory for handling of plasticity has been presented already in \cite{cg1988}. \cite{muller2004point} presents a method for modeling and animating of elastic and plastic objects in realtime using point based animation but it is not widely used.

Goal of this work is to find method that can be used in widely used physics engines to make simulation more realistic
by connecting rigid objects by  spring based constraints which have additional properties for handling plasticity.
Main feature is that plastic deformation takes place if force or moment exceeds given limit and deformation absorbs energy.
This is done by reusing motor code that is already available in most physics engines.

Methods presented in this work can be used in gaming industry to provide more realistic simulations without significant extra work. 
This kind of metodology also opens quite large area of combining old structural analysis methods to modern simulation frameworks.

>>>>

The application of modern computer game techniques enables the description of complex dynamic systems such as military vehicles with a high level of detail while still solving the equations in real-time. Film production and war games, in particular, is a key area that have benefited from simulation technology. In practice, games are often accomplished using an open-source platform such like ODE - \cite{ode}, \bullet\ - \cite{bullet} and Box2D - \cite{box2d}.

Computational methods used in physics engines are divided to modules that handle collision detection and contact description and modules that handle solution of equations in real-time. Equations need to be solved can further be subdivided to be associated to motion, constraints and collisions. Velocity-based formulation is typically used in constraint based rigid body simulation frameworks as collisions cannot be handled easily in acceleration based formulations. Friction is typically taken into account and mechanical joints are handled by constraint equations. Detailed description of various components can be found in e.g. \cite{erleben.thesis}.

Plasticity is not typically taken into account in gaming solutions. Breaking of various objects typically takes place based on collision or impulse. Nevertheless, breaking of steel or reinforced concrete structures using this approach is not appropriate making a simulation to look unrealistic. Theory for handling of plasticity has been presented already in \cite{cg1988}. \cite{muller2004point} and \cite{muller2005meshless} present a method for modeling and animating of elastic and plastic objects in real-time using point based animation. This approach is not been widely used in simulation applications.  On major issue is collision handling of deformable objects.

This study will introduce an approach to account plastic deformation in game applications.   In the introduced method, the plastic deformation takes place if force or moment exceeds given limit, deformation absorbs energy and joint breaks if plastic capacity is exceeced. The approach is based on using joint motors to model plasticity. \cite[p.~90]{erleben.thesis} suggests similar method for modelling friction in joints. Adjacent objects are connected by motors. Motor power production limits are estimated based on plastic section modulus. Joint breaking is accounted by summing plastic deformation and comparing it to predefined material based limit. Elastic part of deformation is modelled by employing spring description which is based on modification of existing constraint in \bullet.

Approach presented in this work can be used in gaming industry to provide more realistic simulations without significant extra work. For gaming purposes presented method works best in scenarios where connected parts are relatively heavy. This allows normal integration timestep to be used without stability issues. This kind of metodology also opens large area of combining old structural analysis methods to modern simulation frameworks.
